---
title: Inhaling Earth
author: Robert McDonnell
date: '2017-05-22'
categories: []
tags: [R]
description: ''
featured: ''
featuredalt: ''
featuredpath: ''
linktitle: ''
draft : true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache.lazy = TRUE, message = FALSE, warning = FALSE, eval = FALSE)
```

After seeing Nadieh Bremer's great [Breathing Earth](http://www.datasketch.es/april/code/nadieh/) infographic, I thought it would be cool to recreate it in R, as you do. Then I saw that it was made from lots of tif files...hmmm. I did some work with those before, ain't doin it again voluntarily, no thanks.   

So then I started thinking about something else that would be (kind of) similar and interesting. I saw the [sf](https://github.com/edzer/sfr) package and its interesting `geom_sf()` recently, and so I thought it would be a nice opportunity to try that out. Given we started with 'breathing' Earth, the natural next step was to think of 'inhaling' Earth! A quick download of some cannabis data from [here](https://data.unodc.org/?lf=1&lng=en) and we're (almost) ready to go, just some cleaning, tidying and merging with the geometry data from the [rnaturalearth](https://github.com/ropenscilabs/rnaturalearth) package. 

```{r}
library(stringi)
library(sf)
library(rnaturalearth)
library(readr)
library(dplyr)
library(ggplot2)

weed <- read_csv("/Users/robert/Downloads/drugs.csv", skip = 2) %>% 
  filter(!is.na(Year)) %>% 
  select(Region, Country, Rate = Best, Year) %>% 
  mutate(Year = ifelse(Year == "2013-2014", "2014", ifelse(
    Year == "2014/15", "2014", ifelse(
      Year == "2013/14", "2014", ifelse(
        Year == "2012/13", "2012", Year)))),
    Year = paste0(Year, "-01-01"), 
    Year = lubridate::parse_date_time(Year, "Ymd"),
    Country = stri_trans_general(Country, "Latin-ASCII"),
    Country = case_when(
      .$Country == "Venezuela (Bolivarian Republic of)" ~ "Venezuela",
      grepl("United Kingdom ", .$Country) ~ "United Kingdom",
      .$Country == "The former Yugoslav Republic of Macedonia" ~ "Macedonia",
      .$Country == "Taiwan Province of China" ~ "Taiwan",
      .$Country == "Russian Federation" ~ "Russia",
      .$Country == "Lao People's Democratic Republic" ~ "Laos",
      grepl("China,", .$Country) ~ "China",
      .$Country == "Bolivia (Plurinational State of)" ~ "Bolivia",
      TRUE ~ .$Country),
    Period = ifelse(Year < "2006-01-01", "2000-2005", ifelse(
      Year > "2010-01-01", "2010-2015", "2006-2010")))

globe <- ne_countries(scale = 110, type = "countries", returnclass =  "sf") %>% 
  select(Country = admin, formal = formal_en, geometry)

two <- full_join(weed, globe) %>% st_as_sf() %>%
  filter(!is.na(formal))

pp <- ggplot(two, aes(frame = Period)) + geom_sf(aes(fill = Rate)) +
  scale_fill_continuous(trans = "reverse", low = "dodgerblue4", high = "white") +
  theme_minimal()

gganimate::gg_animate(pp, interval = 4)

```


Hmm, not a good candidate for animations...


How about booze?


```{r}
'%ni%' <- Negate('%in%')

booze <- read_csv("/Users/robert/Downloads/UNdata_Export_20170523_022116370.csv") %>% 
  select(-Nature, Country = `Reference Area`) %>% 
  filter(Country %ni% c("World", "Developing regions",
                      "Caucasus and Central Asia", "Developed regions",
                      "Eastern Asia", "Landlocked developing countries",
                      "Latin America and the Caribbean", 
                      "Least developed countries", "Oceania",
                      "Northern Africa", "South-eastern Asia",
                      "Small island developing States", "Sub-Saharan Africa",
                      "Western Asia", "Southern Asia")) %>% 
  mutate(Country = case_when(
      .$Country == "Venezuela (Bolivarian Republic of)" ~ "Venezuela",
      grepl("United Kingdom ", .$Country) ~ "United Kingdom",
      .$Country == "The former Yugoslav Republic of Macedonia" ~ "Macedonia",
      .$Country == "Taiwan Province of China" ~ "Taiwan",
      .$Country == "Russian Federation" ~ "Russia",
      .$Country == "Democratic People's Republic of Korea" ~ "North Korea",
      .$Country == "Lao People's Democratic Republic" ~ "Laos",
      .$Country == "Czechia" ~ "Czech Republic",
      .$Country == "Timor-Leste" ~ "East Timor",
      .$Country == "Viet Nam" ~ "Vietnam",
      .$Country == "Brunei Darussalam" ~ "Brunei",
      .$Country == "Bahamas" ~ "The Bahamas",
      .$Country == "Côte d'Ivoire"  ~ "Ivory Coast",
      .$Country == "Guinea-Bissau"  ~ "Guinea Bissau",
      .$Country == "Syrian Arab Republic"  ~ "Syria",
      .$Country == "Iran (Islamic Republic of)"  ~ "Iran",
      .$Country == "Republic of Korea"  ~ "South Korea",
      .$Country == "Congo"  ~ "Republic of Congo",
      .$Country == "Serbia"  ~ "Republic of Serbia",
      .$Country == "Moldova"  ~ "Republic of Moldova",
      .$Country == "Bolivia (Plurinational State of)" ~ "Bolivia",
      TRUE ~ .$Country)) %>% 
  full_join(globe) %>% filter(!is.na(formal))
  


b <- ggplot(booze, aes(frame = `Time Period`)) + geom_sf(aes(fill = Value)) +
  scale_fill_continuous(trans = "reverse", low = "firebrick", high = "white") +
  theme_minimal() + guides(fill = guide_legend(title="Litres of \nPure Alcohol"))

gganimate::gg_animate(b, interval = 4)
```

That works! Nice. Ok, it's pretty simple, but given the hellish wrangling involved with some spatial polygon sets and administrative unit geographical data, I'm really impressed with how easy `geom_sf()` was to use.  





Data from [here](http://data.un.org/Data.aspx?q=alcohol&d=SDGs&f=series%3aSH_ALC_CONSPT). 