---
title: Avoiding the tiresome "training" & "test" data split
author: Robert Myles McDonnell
date: '2018-09-03'
slug: avoiding-the-tiresome-training-test-data-split
categories: [R, destructuring]
tags: []
description: ''
featured: ''
featuredalt: ''
featuredpath: ''
linktitle: ''
---

I really don't like splitting data into 'train' and 'test'. I don't mean that I'm against the idea of it, though you could say it's a waste of data that could be used to better your model, but I mean that actual assignment in R of 'train' and 'test'. I always liked destructuring in Python, and I like it a lot in 2018 JavaScript, so when I remembered that the [zeallot](https://github.com/r-lib/zeallot/blob/master/vignettes/unpacking-assignment.Rmd) package has it, I thought it would be a good opportunity to see how that could fit in a [tidy modelling](https://github.com/tidymodels) pipeline. Much to my delight, one little helper function later, it works perfectly. 

Loading packages and data:  

```{r message=FALSE, warning=FALSE}
library(recipes); library(rsample); library(tidyverse); library(zeallot)

# load some data:
data("credit_data")
``` 

Next comes our little helper function: 

```{r}
m_bake <- function(recipe_object, data){
  cd <- initial_split(credit_data)
  tr <- training(cd)
  te <- testing(cd)
  x1 <- bake(recipe_object, tr)
  x2 <- bake(recipe_object, te)
  return(list(x1, x2))
}
```

Now you get a nice clean pipeline with a train/test split as a result, using zeallot's great `%->%` operator:  
```{r}
recipe(Status ~ ., data = credit_data) %>% 
  step_knnimpute(all_predictors()) %>% 
  step_center(all_numeric()) %>% 
  step_dummy(-all_numeric()) %>% 
  prep() %>% 
  m_bake(data = credit_data) %->% c(train, test) %->% c(train, test)

ls()

head(train); head(test)
```

Lovely.  

(Ok, I still had to split it (!!), but once you write this function, you can just call this.)
