---
title: Cold Oktober
author: Robert Myles McDonnell
date: '2018-10-27'
slug: cold-oktober
categories: [R, weather, D3]
tags: []
description: 'Am I cold or am I imagining it?'
featured: ''
featuredalt: ''
featuredpath: ''
linktitle: ''
---

Moving back to Ireland from Brazil has me feeling cold. This got me thinking that seems like an unusually cold October, but, since I just moved here, I may not be a reliable guide. I was wondering whether I was a) correct, or b) a weakling, unable to withstand temperatures I would have previously considered normal. Well, I did what every data scientist would do in such a situation -- I found data, cleaned it up and made graphs. I was able to scrape some recent data from [met eireann's](http://archive.met.ie/climate/monthly-data.asp?Num=518) historical site and download historical data from the World Bank's [Climate Change Knowledge Portal](http://sdwebx.worldbank.org/climateportal/index.cfm?page=country_historical_climate&ThisCCode=IRL), both of which needed some cleanup:  

```{r dataprep, message=FALSE, warning=FALSE}
library(tidyverse); library(glue);library(rvest)
library(readxl); library(lubridate);
library(r2d3)

source("https://raw.githubusercontent.com/RobertMyles/r_utils/master/char_from_copy.R")
source("https://raw.githubusercontent.com/RobertMyles/r_utils/master/month_changer.R")


tabl <- read_html('http://archive.met.ie/climate/monthly-data.asp?Num=518') %>% 
  xml_nodes('#climate-report > table:nth-child(5)') %>% 
  html_table() %>% .[[1]] %>% 
  magrittr::set_colnames(
    value = char_from_copy(
      "Year Jan Feb Mar Apr May Jun Jul Aug Sep Oct Nov Dec Annual"
      )) %>% 
  slice(-1) %>% 
  mutate_all(as.numeric) %>% 
  slice(-5) %>% select(-Annual) %>% 
  gather("Month", "temp", -Year) %>% 
  month_changer("Month") %>% 
  mutate(Day = "01",
         date = glue('{Year}-{Month}-{Day}')) %>% 
  select(-c(Year, Month, Day)) %>% 
  filter(year(parse_date_time(date, 'ymd')) > 2015)


temps <- read_excel('~/Downloads/tas_1901_2015.xls') %>% 
  rename("temp" = "tas") %>% 
  mutate(Day = "01",
         date = glue('{Year}-{Month}-{Day}')) %>% 
  select(-c(Country, Year, Month, Day, ISO3, ISO2)) %>% 
  bind_rows(tabl) %>% 
  filter(!is.na(temp)) %>% 
  arrange(parse_date_time(date, 'ymd'))

```

Combining these gives us this table:  

```{r}
knitr::kable(head(temps))
```

```{r}
knitr::kable(tail(temps))
```

<script src="https://cdnjs.cloudflare.com/ajax/libs/d3-annotation/2.3.1/d3-annotation.js"></script>

```{d3 echo = FALSE, data=temps}
margin = {top: 10, bottom: 20, left: 30, right: 5}
w = 670
h = 400

d2 = data
d2.forEach(d => d.date = new Date(d.date))
d2.forEach(d => d.mean = d3.mean(d2, d => d.temp))

x = d3.scaleTime()
    .domain(d3.extent(d2, d => d.date))
    .range([margin.left, w - margin.right])

y = d3.scaleLinear()
    .domain([0, d3.max(d2, d => d.temp)]).nice()
    .range([h - margin.bottom, margin.top])

xAxis = g => g
    .attr("transform", `translate(0,${h - margin.bottom})`)
    .call(d3.axisBottom(x).ticks(w/60).tickSizeOuter(0))

yAxis = g => g
    .attr("transform", `translate(${margin.left},0)`)
    .call(d3.axisLeft(y))
    .call(g => g.select(".domain").remove())
    .call(g => g.select(".tick:last-of-type text").clone()
        .attr("x", 3)
        .attr("text-anchor", "start")
        .attr("font-weight", "bold")
        .text(d2.y))

line = d3.line()
    .x(d => x(d.date))
    .y(d => y(d.temp))

av_line = d3.line()
    .x(d => x(d.date))
    .y(d => y(d.mean))

svg.append("g").call(xAxis);

svg.append("g").call(yAxis);
  
svg.append("path")
    .datum(d2)
    .attr("fill", "none")
    .attr("stroke", "steelblue")
    .attr("stroke-width", 1.5)
    .attr("stroke-linejoin", "round")
    .attr("stroke-linecap", "round")
    .attr("d", line)

svg.append("path")
    .datum(d2)
    .attr("fill", "none")
    .attr("stroke", "#B84A62")
    .attr("stroke-width", 1.5)
    .attr("stroke-linejoin", "round")
    .attr("stroke-linecap", "round")
    .attr("d", av_line);

svg.append("text")
    .attr("x", 550)
    .attr("y", 360)
    .attr("dy", ".35em")
    .style("fill", "#B84A62")
    .text("Mean: 9.2 degrees");

svg.append("line")
    .attr("x1", 550)
    .attr("x2", 470)
    .attr("y1", 345)
    .attr("y2", 190)
    .style("stroke", "#B84A62");
```

